\mytitle

### 2015-08-04 MSc_diss.tex
Updates in mean RONO2 concentrations, clean up literature review

\documentclass[11pt,a4paper]{article}
% Encoding and language
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
% Colors
\usepackage[usenames,dvipsnames]{color}
\usepackage{colortbl}
% Paper size and margins
%\usepackage{rotating}
\usepackage{pdflscape}
%\usepackage[a4paper,showframe]{geometry}
\usepackage{vmargin}
\setmarginsrb{2cm}{2cm}{2cm}{2cm}{0cm}{0cm}{0cm}{1.5cm}
\usepackage{pdflscape}
\usepackage{afterpage}
% Indenting
\usepackage{indentfirst}
%\setlength{\parindent}{1cm}
\setlength{\parskip}{0.5cm}
% Symbols
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
% Equations
\usepackage{eulervm} % upright font in eqns (Palatino)
\usepackage{cool}
\usepackage{mathtools} % for arrows
\renewcommand{\theequation}{R\arabic{equation}}
% References
\usepackage{natbib}
\bibliographystyle{apalike}
% Hyperlinks and labels
%\usepackage{showkeys}
\usepackage[linktocpage=true,plainpages=false,pdfpagelabels=false]{hyperref}
\definecolor{citecolor}{rgb}{0.2,0.7,0.9}
\definecolor{urlcolor}{rgb}{0,0,1}
\hypersetup{
    colorlinks, linkcolor={Blue},
    citecolor={Blue}, urlcolor={Black}
}
% Figures
\usepackage{subcaption}
\usepackage{alphalph}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{positioning}
\usetikzlibrary{calc}
\usepackage{epstopdf}
% Tikz styles
\tikzstyle{reg} = [rounded rectangle, text width=2cm, fill=black!5, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]
\tikzstyle{rad} = [rounded rectangle, draw=none, text width=2cm, fill=red!10, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]
\tikzstyle{long} = [rounded rectangle, draw=none, text width=3.5cm, fill=black!5, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]   
\tikzstyle{add} = [scale=1.5, draw=none,fill=none,align=center]
\tikzstyle{emp} = [draw=none,fill=none]
\tikzstyle{ar} = [->,draw=black!70,line width=2]
\tikzstyle{ln} = [-,draw=black!70,line width=2]
% Title etc
\newcommand{\mytitle}{Investigation of the relationship between tropospheric ozone production efficiency and carbon bond emissions}
\newcommand{\authorname}{Maria Zamyatina}
\newcommand{\authornumber}{Student No: 100106685}
\newcommand{\degname}{Master of Science}
\newcommand{\univname}{University of East Anglia}
\newcommand{\univaddr}{University Plain \\ Norwich NR47TJ UK}
\newcommand{\depname}{School of Environmental Sciences}
\newcommand{\mydate}{\the\year}
%% Remove author and date from the maketitle command
%\makeatletter
%\renewcommand{\@maketitle}{
%\newpage
% \null
% \vskip 2em%
% \begin{center}%
%  {\LARGE \@title \par}%
% \end{center}%
% \par} \makeatother

@@

During/After one of these investigations, namely the SAGA campaign, \citep{Atlas1993} proposed by that the short-chained alkyl nitrates ($C_1$-$C_3$) can be produced not only photochemically through the oxidation of their parent hydrocarbons \citep{Roberts1990}, but also emitted directly from equatorial oceans. This has since been
confirmed by field measurements \citep{Chuck2002,Blake2003,Dalh2005}.

There have been a number of studies of the latitudial and seasonal variations of alkyl nitrates in marine air (Blake et al., 2003), in the Arctic (Swanson et al., 2003) and in the
Antarctic (Beyersdorf et al., 2010)
\subsubsection{Observational studies}
Anthropogenic!!! Neuman
Source characterization and photochemical processes in transported air masses have been analyzed by examining RONO2 formation from anthropogenic emissions of alkane precursors [Simpson et al., 2003; Madronich, 2006; Simpson et al., 2006; Reeves et al., 2007; Worton
et al., 2010]

\citep{Sommariva2008}
[Alkyl nitrates typically constitute only a small
fraction of total NOy <10\%, even though they can be a much larger component in the equatorial marine boundary layer (Blake et al., 2003; Simpson et al., 2006 and references therein). Due to their long lifetime (41 week) alkyl nitrates can transport NOx to remote regions, with potential impacts on regional O3 formation (e.g., Murphy et al., 2006; Simpson et al., 2006)]

@@

\citep{Stroud2001} plane+colorado, plane+North Atlantic
During both campaigns alkyl nitrates were measured aboard the NOAA Orion WP-3 aircraft with a vertical range from the boundary layer to the midtroposphere (7 km) [Hiibler et al.,1998]. During STERAO 1996 the aircraft was based at Buckley Air Force Base in Denver, Colorado. Flights were conducted from June 13, 1996, to July 16, 1996, over the northeastern plains of Colorado to study how thunderstorms affect the chemical composition of the atmosphere.
Alkyl nitrates were measured over the northeastern plains of Colorado during the Stratospheric Tropospheric Exchange: Radiation, Aerosols and Ozone (STERAO) 1996 campaign and over the North Atlantic during the North Atlantic Regional Experiment (NARE) 1997. A comparison of alkyl nitrate to hydrocarbon ratios with a sequential reaction scheme model suggests our understanding of C4-Cs alkyl nitrate photochemistry is adequate to reproduce the observations. Air masses with larger ratios of alkyl nitrate to hydrocarbons were observed during NARE 1997 than STERAO 1996, consistent with the NARE 1997 measurements being in the more remote North Atlantic. These results suggest that information related to the photochemical age of an air  mass can be derived from alkyl nitrate to hydrocarbon ratios.

\citep{Worton2010} TORCH
Alkyl nitrates (C1eC5) were measured at two sites (near urban and rural) in southeast England during the Tropospheric Organic Chemistry Experiment (TORCH). Methyl nitrate was the dominant species during both campaigns accounting for on average about one third of the total measured alkyl nitrates. High mixing ratios (>50 pptv) and variability of methyl nitrate were observed at the near urban site (TORCH1) that were not seen at the rural site (TORCH2) and which could not be explained by local photochemical production or direct emissions. The diurnal variation of methyl nitrate during TORCH1 showed a morning maximum that would be consistent with nighttime chemistry followed by transport to the surface by boundary layer dynamics. Similarly, elevated morning mixing ratios were also observed during TORCH2 although the magnitudes were much smaller. As a result, methyl nitrate could represent a tracer for nighttime chemistry seen at the ground the following day. At both campaigns, the dominant source of short chain alkyl nitrates and carbonyl precursor radicals (?C4) were from decomposition of larger compounds. The magnitude of the source increased with decreasing carbon number consistent with increasing total precursor abundance. Non-photochemical emissions of acetaldehyde and acetone could not be accounted for by automobile exhaust emissions alone and indicated that other direct sources are likely important in this environment.
Alkyl nitrate measurements were made at two sites (near urban
and rural) as part of the TORCH experiment. Observed mixing ratios
were higher (factor of ?2) during TORCH1, especially during the identified heatwave period. Methyl nitrate accounted for w30\% of the total measured C1eC5 species and during TORCH1 was often present at mixing ratios > 50 pptv. The diurnal profile of methyl nitrate showed an early morning maximum that we attributed to nighttime production from NO3 chemistry in the residual layer followed by transportation to the surface by boundary layer dynamics. Further work is needed to evaluate our hypothesis of a nighttime source of methyl nitrate but if proven, methyl nitrate could represent an interesting tracer of nighttime chemistry aloft. The predominant photochemical source of the short chain alkyl nitrates and carbonyl compounds was from the photochemical oxidation and decomposition of longer chain compounds as opposed to the oxidation of the parent hydrocarbons similar to previous studies. The magnitude of this source shows a preference for the shortest chain species consistent with the increasing number of precursor compounds identified in the MCM. Primary sources of acetaldehyde and acetone were also significant during TORCH1 and could not be accounted for by automobile exhaust emissions alone. Further work to identify and quantify emissions from additional direct sources is necessary if appropriate control strategies are to be implemented.

\citep{deKock1994}
Air samples were collected at Cape Recife (34 1' S 25 40'E) situated on the south-east coast of Southern Afi'ica (Figure 1). The suburban, urban and industrial areas of Port Elizabeth lie approximately 10-40 kin north to north-west of the sampling site.
Mean sumANs 12.7.
Variations in absolute and relative concentrations of the alkyl nitrates could be related to changes in meteorological conditions. It is considered that the differences observed in chemical composition may represent different sources of trace gas emissions or different degrees of photochemical processing of similar sources. The meas~ed levels of the alkyl nitrates suggest long-range transport to play a major role in regulating the distribution of alkyl nitrates in the troposphere.

\citep{Beyersdorf2010}
South Pole, other Antarctic, ANs only

\subsubsection*{Oceanic sources}
\citep{Newland2013}
[An oceanic source for C1 – C3 alkyl nitrates was first postulated by Atlas et al. (1993) during the SAGA campaign in which it was observed that elevated mixing ratios of methyl and ethyl nitrates correlated well with equatorial regions of high productivity. This has since been confirmed by field measurements (e.g. Chuck et al., 2002; Blake et al., 2003; Dahl et al., 2005). Chuck et al. (2002) estimate an annual emission flux from equatorial Atlantic and Pacific Ocean regions of 230 and 68 Gg yr-1 for methyl and ethyl nitrate respectively. The oceanic source of more or equal C4 alkyl nitrates is insignificant compared to anthropogenic sources (Blake et al., 2003).
The production mechanism of the alkyl nitrates is thought to be by the reaction ROO + NO in the aqueous phase with the peroxy radical coming from the photolysis of coloured dissolved organic matter (CDOM) and the NO coming from the photolysis of nitrite (Dahl et al., 2003). Direct emission from algae (Chuck et al., 2002) or from bacteria at depth (Dahl et al., 2007) has also been postulated.]

\citep{Atlas1993}
Alkyl Nitrates, Nonmethane Hydrocarbons, and Halocarbon Gases Over the Equatorial Pacific Ocean During Saga 3
February March 1990
The cruise covered a region of the equatorial Pacific Ocean from 15N to 10S latitude and 144 to 165W longitude. On this cruise we collected samples for the measurement of alkyl nitrates (RONO2) (C2-C5), nonmethane hydrocarbons (NMHC) and several halocarbon gases.
Results:
- Alkyl nitrates were found at levels higher than predicted from steady state calculations based on measured mixing ratios of hydrocarbons and NO. The measured levels of RONO2 suggest long-range transport as one mechanism contributing to elevated concentrations of alkyl nitrates in the remote troposphere. However, the distributions of C2 and C3 alkyl nitrates over the equator were  similar to the organobromine gases. This distribution suggests
a possible oceanic source for alkyl nitrates to the atmosphere.
Model: 1D 41 constituents incl. 3 ANs

\citep{Chuck2002} September, October
Measurements of methyl and ethyl nitrate in seawater and air samples along two Atlantic Ocean transects provide the FIRST direct evidence for an oceanic source of these compounds. A simple box model calculation suggests that the equatorial source could be a major component of the local atmospheric alkyl nitrate budget.
In the remote atmosphere, the concentration of methyl nitrate (MeONO2) is expected to be low because of low concentrations of NOx and a branching ratio of less than 1\% for the formation of MeONO2
from the reaction of alkylperoxy radicals with NO (8). However, unexpect- edly high levels of RONO2 in marine air and observations of a strong positive correlation between ethyl and propyl nitrate and the marine biogenically produced halocarbon, bromoform, in the marine boundary layer over the equatorial Pacific Ocean led Atlas et al. to propose an oceanic source for alkyl nitrates in the atmo- sphere (9). Subsequent atmospheric sampling campaigns, mostly conducted over the remote Pacific Ocean, have provided yet more evidence to support the hypothesis of an oceanic source (2, 10), with the contribution of alkyl nitrates to the NOy
sum reaching 80\% in equatorial and high-latitude regions over the South Pacific (Blake, 1999). Although there have been no pub- lished measurements confirming the oceanic emission of these species, other studies have also invoked a seawater source for methyl, eth- yl, and propyl nitrates (11, 12). The supersaturation of surface waters over large areas of the Atlantic provides the first direct evidence for an oceanic source of MeONO2 and EtONO2. dominant source region is found between the equator and 40S. distinct mechanisms may be respon- sible for the production of methyl and ethyl nitrate, with different water types supporting different mechanisms. Methyl nitrate generally showed profiles more typically associated with algal production, with decreasing concentrations be- low the chlorophyll maximum. In contrast, in waters more strongly affected by continental influences, such as increased loading of dis- solved organic matter, the alkyl nitrates showed maximum concentrations in surface waters, which may suggest that photochemical production is important in shelf and coastal regions.

\citep{Blake2003} oceanic+urban!!!
Observations: concentration distributions of C1-C4 alkyl nitrates observed during the NASA airborne campaigns Pacific Exploratory Mission (PEM) -Tropics A (September– October 1996) and PEM-Tropics B (March–April 1999). The total geographic range for PEM-Tropics A was 45?N–72?S latitude and 153?E–75?W longitude, and for PEM- Tropics B was 40?N–36?S latitude and 149?E–75?W longitude. We observed high methyl nitrate (MeONO2,CH3ONO2) mixing ratios (approximately 50 pptv) at low altitudes in a latitude band between 8?Nto13?S stretching across the equatorial Pacific, illustrating the oceanic source of MeONO2. This source may be associated with the high-nutrient, low-chlorophyll character of equatorial Pacific waters. We discuss MeONO2 and ethyl nitrate (EtONO2,C2H5ONO2), whose abundance is dominated by equatorial oceanic sources, 2-Propyl nitrate (2-PrONO2,2-C3H7ONO2), which has significant oceanic and northern hemispheric (NH) sources associated with urban/ industrial hydrocarbon emissions, and 2-butyl nitrate (2-BuONO2 2-C4H8ONO2), which has mostly NH (northern hemisphere) sources.
the ocean as the major source for MeONO2
Ethyl nitrate has a dominant oceanic source in equatorial regions, but at mid NH latitudes a small additional source is associated with urban/industrial emissions.

\citep{Reeves2007} ocean+urban+fire
This paper is based on alkyl nitrate measurements made over the North Atlantic as part of the International Consortium for Research on Atmospheric Transport and Transformation (ICARTT). The focus is on the analysis of air samples collected on the UK BAe-146 aircraft during the Intercontinental Transport of Ozone and Precursors (ITOP) project, but air samples collected on board the NASA DC-8 and NOAA WP-3D aircraft as part of a Lagrangian experiment are also used. The ratios between the alkyl nitrates and their parent hydrocarbons are compared with those expected from chemical theory. Further, a box model is run to investigate the temporal evolution of the alkyl nitrates in three Lagrangian case studies and compared to observations.
The general agreement between the alkyl nitrate data and photochemical theory suggests that during the first few days of transport from the source region, photochemical production of alkyl nitrates, and thus ozone, had taken place.
Concentrations of methyl, ethyl, 1-propyl, 2-propyl, 2-butyl, 2-methyl-3-butyl and the sum of 2- and 3-pentyl nitrates (hereafter referred to as 2 + 3-pentyl nitrate) are presented.
The alkyl nitrate concentrations observed during
ITOP are typical of those found a few days downwind of precursor source regions. Although there is some evidence for an oceanic source of methyl nitrate, for the most part the air samples collected during ITOP do not appear to be strongly influenced by oceanic sources of the alkyl nitrates. Elevated concentrations were observed in air masses which have been influenced by emissions from the N.E. United States and from Alaskan fires. The large variations in ratios between the C5 and C2-C4 nitrates suggest that emission ratios for the parent hydrocarbons vary between different regions of the N.E. United States. [49] In general the alkyl nitrate data agrees well with photochemical theory. This suggests that at least during the first few days of transport from the source region photo- chemical production of alkyl nitrates had taken place. This implies that ozone production had also occurred. There also appears to be a widespread common source of ethyl nitrate and 1-propyl nitrate other than from their parent hydro- carbons. This might be from longer-chain hydrocarbons, possibly pentanes. [50] The observations in the more photochemically pro- cessed air masses are consistent with the concentration of NO dropping to a point where the rates of the alkyl nitrate production reactions no longer dominate the peroxy radical self/cross reactions. However, in these low levels of NO, where the balance between ozone production and destruc- tion is highly sensitive to NOX, the alkyl nitrates might contribute to maintaining the ozone levels in air as it undergoes long-range transport. [51] The observed patterns of the alkyl nitrates also suggest that the rates of photochemical processing in the Alaskan smoke plumes were small and that the rates of ozone formation in these air masses were low.

\citep{Simpson2003} airborne too
TRACE-P samples that did not show influence from marine alkyl nitrate sources were used to investigate photochemical alkyl nitrate evolution. Relative to 2-butyl nitrate/n-butane, the measured ratios of ethyl nitrate/ethane and 2-propyl nitrate/propane showed notable deviations from modeled values based on laboratory kinetic data, suggesting additional Asian sources of their alkyl peroxy radical precursors. By contrast, the measured ratios of 1-propyl-, 2-pentyl-, and 3-pentyl nitrate to their respective parent hydrocarbons were fairly close to modeled values.

\citep{Roberts1998} marine MBL+urban ANs
The measurements described herein were made as part of the
Organic nitrates are a class of product species that contain
North Atlantic Regional Experiment (NARE) during August and September 1993. These measurements consisted of NO, NO 2, particulate nitrate, HNO3, PAN, PPN, RONO 2, NOy, 0 3, CO, SO 2, and C2-C6 hydrocarbons at Chebogue Point. This site was ideally located to study the evolution of photochemically processed air masses in the marine boundary layer (MBL). Chebogue Point is approximately 500 km downwind from the eastern sea coast of the United States, and is often impacted by transport of polluted air from the continental source regions.
Northwest Territories, Canada, have been made in the winter to
The correlations between organic nitrates and CO or O3 have been presented by Roberts et al. [1996]. It has been established that CO is a good tracer of continentally derived pollution at this site [Parrish et al., 1993]. This continentally derived pollution contains hydrocarbons (HC) and NO x, the precursors of organic nitrates and ozone, and serves to link the sources of CO with the sources of organic nitrates and ozone. There are distinct differences in the lifetimes of CO, O3 , and organic nitrates, and the factors that they are dependent on. For these reasons, organic nitrates correlate to varying degrees with CO and O3.
Measurements of the organic nitrates; PAN, PPN, methyl-,
ethyl-, 1-propyl-, 2-propyl-, 2-butyl, and 3-pentyl nitrates were made at Chebogue Point, Nova Scotia, during the 1993 NARE campaign. The relationships of these compounds with other measures of continental pollution, for example, CO and 03, are consistent with our view of these compounds as products of HC-NO x photochemistry.
Comparisons of the ratios with that of [2-butyl nitrate]/[butane] were used to test whether the actual data fit the line predicted from rate constants, branching ratios, photolysis rates, and assumed [OH].
Abstract. Measurements of six alkyl nitrates, methyl-, ethyl-, 1-propyl-, 2-propyl-, 2-butyl, and 3-pentyl nitrates, and two peroxycarboxylic nitric anhydrides, peroxyacetic nitric anhydride
(PAN) and peroxypropionic nitric anhydride (PPN) were made
at Chebogue Point, Nova Scotia, during the 1993 North Atlantic
Regional Experiment campaign. The correlation of alkyl nitrates
(RONO2) with carbon monoxide makes it clear that the chemistry at this site is being driven by continental emissions, since RONO2 compounds are known to be by-products of HC-NOx photochemistry.
Comparison of the ratios of alkyl nitrates to their parent hydrocarbons to that of 2-butyl nitrate/butane showed significant
deviations from trends predicted from rate constants, branching
ratios, and loss rates. For ethyl nitrate this deviation is likely due to a combination of many additional pathways  o the intermediate
peroxy radical from the decomposition of larger alkoxy radicals.
The propyl nitrates showed trends that differed from one another; 2-propyl nitrate was a factor of 2 to 3 higher than predicted, while 1-propyl nitrate was a factor of 8 to 10 higher than predicted. The trend of 3-pentyl nitrate/pentane was relatively close (about
50\% lower) to the predicted trend.

\subsubsection*{Biomass burning source}
\citep{Newland2013}
[Alkyl nitrates also have a biomass burning source (Friedli et al., 2001; Simpson et al., 2002). Simpson et al. (2002) report emissions of C1 – C4 alkyl nitrates from savanna fires in Australia, with methyl nitrate being primarily emitted during the ’flaming’ stage and ethyl – butyl nitrates mainly emitted during the ‘smouldering’ stage. They estimate global emissions from this source to be on the order of 8 Gg yr-1]
\citep{Simpson2002}
We report the first observations of the emission of five C1–C4 alkyl nitrates (methyl-, ethyl-, n-propyl-, i-propyl-, and 2-butyl nitrate) from savanna burning. Average alkyl nitrate mixing ratios in the immediate vicinity of three bushfires in Northern Australia were 47–122 times higher than local background mixing ratios. These are the highest alkyl nitrate mixing ratios we have ever detected, with maximum mixing ratios exceeding 3 ppbv for methyl nitrate. Methyl nitrate dominated the alkyl nitrate emissions during the flaming stage of savanna burning, whereas C2–C4 alkyl nitrates were mostly emitted during the smoldering stage. To explain the formation of alkyl nitrates from biomass burning, we propose a reaction mechanism involving the combination of reactive radicals at high temperature. Bearing in mind the uncertainties associated with extrapolating small data sets to much larger scales, alkyl nitrate emissions from global savanna burning are estimated to be on the order of 8 Gg/yr.

\citep{Chuck2002}
Use of the mean sea-to-air fluxes for 0 to 40S (Table 1) and measured atmospheric concentrations results in a lifetime of be- tween 4.5 and 25 days for MeONO2 and between 5 and 10 days for EtONO2 (22). Literature estimates of the atmospheric life- time of MeONO2, based on known atmospheric destruction processes, are poorly con- strained, ranging from 6 to 29 days (23, 24). For EtONO2
, the atmospheric lifetime with respect to photodissociation at the equator and at 40S has been estimated to be 7 and 12 days, respectively, for October (25).

\subsubsection*{Global distribution}
\citep{Newland2013}
[There have been a number of studies of the latitudinal and seasonal variations of alkyl nitrates in marine air (Blake et al., 2003), in the Arctic (Swanson et al., 2003) and in the Antarctic (Beyersdorf et al., 2010).
Blake et al. (2003) report extensive measurements of C1 – C4 alkyl nitrates from the Pacific between 45N and 72S during the PEM-Tropics A and B projects. This study observed an equatorial band between about 8N and 13S in which the oceans are a source of alkyl nitrates. The highest mean methyl and ethyl nitrate measurements were found in this region. The highest mixing ratios of 2-propyl nitrate were found in the most northerly regions of the study (i.e. 40-45N) but it also showed significantly elevated mixing ratios in the same equatorial band as the C1 – C2 nitrates. The high mixing ratios in the northern mid-latitudes are explained by production from anthropogenic polluted outflow. 2-butyl nitrate showed almost no enhancement in equatorial regions and again had the highest mixing ratios in the most northerly regions of the study.
Swanson et al. (2003) and Beyersdorf et al. (2010) observe a strong seasonality of alkyl nitrate mixing ratios in the Arctic and Antarctic respectively. Polar alkyl nitrate mixing ratios show a similar seasonality to the alkanes, with a spring-time peak just before polar sunrise and lows in the summer time. Winter time mixing ratios of 2-propyl and 2-butyl nitrate are about 10 ppt, while methyl and ethyl nitrate are about 5 – 6 ppt. In the summer 2-propyl and 2-butyl nitrate mixing ratios fall close to zero, while methyl and ethyl nitrate mixing ratios fall to about 3 ppt. The amplitude of the seasonal cycle reflects the reaction rate with OH, i.e. it increases with increasing carbon number.]

>>>>

\documentclass[11pt,a4paper]{article}
% Encoding and language
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
% Colors
\usepackage[usenames,dvipsnames]{color}
\usepackage{colortbl}
% Paper size and margins
%\usepackage{rotating}
\usepackage{pdflscape}
%\usepackage[a4paper,showframe]{geometry}
\usepackage{vmargin}
\setmarginsrb{2cm}{2cm}{2cm}{2cm}{0cm}{0cm}{0cm}{1.5cm}
\usepackage{pdflscape}
\usepackage{afterpage}
% Indenting
\usepackage{indentfirst}
%\setlength{\parindent}{1cm}
\setlength{\parskip}{0.5cm}
% Symbols
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
% Equations
\usepackage{eulervm} % upright font in eqns (Palatino)
\usepackage{cool}
\usepackage{mathtools} % for arrows
\renewcommand{\theequation}{R\arabic{equation}}
% References
\usepackage{natbib}
\bibliographystyle{apalike}
% Hyperlinks and labels
%\usepackage{showkeys}
\usepackage[linktocpage=true,plainpages=false,pdfpagelabels=false]{hyperref}
\definecolor{citecolor}{rgb}{0.2,0.7,0.9}
\definecolor{urlcolor}{rgb}{0,0,1}
\hypersetup{
    colorlinks, linkcolor={Blue},
    citecolor={Blue}, urlcolor={Black}
}
% Figures
\usepackage{subcaption}
\usepackage{alphalph}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{positioning}
\usetikzlibrary{calc}
\usepackage{epstopdf}
% Tikz styles
\tikzstyle{reg} = [rounded rectangle, text width=2cm, fill=black!5, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]
\tikzstyle{rad} = [rounded rectangle, draw=none, text width=2cm, fill=red!10, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]
\tikzstyle{long} = [rounded rectangle, draw=none, text width=3.5cm, fill=black!5, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]   
\tikzstyle{add} = [scale=1.5, draw=none,fill=none,align=center]
\tikzstyle{emp} = [draw=none,fill=none]
\tikzstyle{ar} = [->,draw=black!70,line width=2]
\tikzstyle{ln} = [-,draw=black!70,line width=2]
% Title etc
\newcommand{\mytitle}{Investigation of the relationship between tropospheric ozone production efficiency and carbon bond emissions}
\newcommand{\authorname}{Maria Zamyatina}
\newcommand{\authornumber}{Student No: 100106685}
\newcommand{\degname}{Master of Science}
\newcommand{\univname}{University of East Anglia}
\newcommand{\univaddr}{University Plain \\ Norwich NR47TJ UK}
\newcommand{\depname}{School of Environmental Sciences}
\newcommand{\mydate}{\the\year}

\newcommand{\picspath}{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/}
%% Remove author and date from the maketitle command
%\makeatletter
%\renewcommand{\@maketitle}{
%\newpage
% \null
% \vskip 2em%
% \begin{center}%
%  {\LARGE \@title \par}%
% \end{center}%
% \par} \makeatother

@@

During/After one of these investigations, namely the SAGA campaign, \citep{Atlas1993} first postulated that the short-chained alkyl nitrates ($C_1$-$C_3$) can be produced not only photochemically through the oxidation of their parent hydrocarbons \citep{Roberts1990}, but also emitted directly from equatorial oceans. This has since been confirmed by field measurements \citep{Chuck2002,Blake2003,Dahl2005}. \citep{Chuck2002} provided the first direct evidence of that based on detection of supersaturation of surface waters by methyl and ethyl nitrates and their high concentrations in the marine air over equatorial areas (0-40$^{\circ}$) of the Atlantic. Later during two extensive NASA airborne campaigns Pacific Exploratory Mission (PEM)-Tropics A (September-October 1996) and PEM-Tropics B (March-April 1999) \citep{Blake2003} observed high levels of methyl and ethyl nitrates (approximately 50 ppt) in the air at the band between 8$^{\circ}$N-13$^{\circ}$S stretching across the equatorial Pacific, and pointed out that oceanic source of $\geq C_4$ alkyl nitrates is insignificant compared to anthropogenic sources.

Alkyl nitrates also have a biomass burning source which have been first reported by \citep{Simpson2002}. They found that in the immediate vicinity of savanna fires in Northern Australia the average alkyl nitrate mixing ratios were 47–122 times higher than local background mixing ratios, exceeding 3 ppbv for methyl nitrate. Methyl nitrate was primarily emitted during the 'flaming' stage, while $C_2$-$C_4$ ethyl-butyl nitrates dominated during the 'smoldering' stage.

There have been a number of studies of the seasonal variations of alkyl nitrates in the Arctic \citep{Swanson2003} and in the Antarctic \citep{Beyersdorf2010}. Polar alkyl nitrate follow the seasonality of the alkanes displaying a clear late winter maximum and a broad summer minimum.

Owing to the existence of the relationship between alkyl nitrates and their parent hydrocarbons, measurements of their concentrations can provide information on the extent of photochemical processing in the air mass. \cite{Bertman1995} developed a kinetic approach to describe this relationship through the evolution of the ratio of alkyl nitrate to its parent hydrocarbon as a function of time. This approach has been extensively used to study air mass photochemical ageing at several sites, in North America \citep{Bertman1995,Roberts1998}, China \citep{Simpson2006}, the British Isles \citep{Worton2010} and over the Pacific \citep{Simpson2003} and North Atlantic Ocean \citep{Reeves2007,Stroud2001}.

@@

\citep{deKock1994}
Air samples were collected at Cape Recife (34 1' S 25 40'E) situated on the south-east coast of Southern Afirica (Figure 1). The suburban, urban and industrial areas of Port Elizabeth lie approximately 10-40 kin north to north-west of the sampling site.
Mean sumANs 12.7.
Variations in absolute and relative concentrations of the alkyl nitrates could be related to changes in meteorological conditions. It is considered that the differences observed in chemical composition may represent different sources of trace gas emissions or different degrees of photochemical processing of similar sources. The meas~ed levels of the alkyl nitrates suggest long-range transport to play a major role in regulating the distribution of alkyl nitrates in the troposphere.

\subsubsection*{Oceanic sources}
\citep{Reeves2007} ocean+urban+fire
This paper is based on alkyl nitrate measurements made over the North Atlantic as part of the International Consortium for Research on Atmospheric Transport and Transformation (ICARTT). The focus is on the analysis of air samples collected on the UK BAe-146 aircraft during the Intercontinental Transport of Ozone and Precursors (ITOP) project, but air samples collected on board the NASA DC-8 and NOAA WP-3D aircraft as part of a Lagrangian experiment are also used. The ratios between the alkyl nitrates and their parent hydrocarbons are compared with those expected from chemical theory. Further, a box model is run to investigate the temporal evolution of the alkyl nitrates in three Lagrangian case studies and compared to observations.
The general agreement between the alkyl nitrate data and photochemical theory suggests that during the first few days of transport from the source region, photochemical production of alkyl nitrates, and thus ozone, had taken place.
Concentrations of methyl, ethyl, 1-propyl, 2-propyl, 2-butyl, 2-methyl-3-butyl and the sum of 2- and 3-pentyl nitrates (hereafter referred to as 2 + 3-pentyl nitrate) are presented.
The alkyl nitrate concentrations observed during
ITOP are typical of those found a few days downwind of precursor source regions. Although there is some evidence for an oceanic source of methyl nitrate, for the most part the air samples collected during ITOP do not appear to be strongly influenced by oceanic sources of the alkyl nitrates. Elevated concentrations were observed in air masses which have been influenced by emissions from the N.E. United States and from Alaskan fires. The large variations in ratios between the C5 and C2-C4 nitrates suggest that emission ratios for the parent hydrocarbons vary between different regions of the N.E. United States. [49] In general the alkyl nitrate data agrees well with photochemical theory. This suggests that at least during the first few days of transport from the source region photo- chemical production of alkyl nitrates had taken place. This implies that ozone production had also occurred. There also appears to be a widespread common source of ethyl nitrate and 1-propyl nitrate other than from their parent hydro- carbons. This might be from longer-chain hydrocarbons, possibly pentanes. [50] The observations in the more photochemically pro- cessed air masses are consistent with the concentration of NO dropping to a point where the rates of the alkyl nitrate production reactions no longer dominate the peroxy radical self/cross reactions. However, in these low levels of NO, where the balance between ozone production and destruc- tion is highly sensitive to NOX, the alkyl nitrates might contribute to maintaining the ozone levels in air as it undergoes long-range transport. [51] The observed patterns of the alkyl nitrates also suggest that the rates of photochemical processing in the Alaskan smoke plumes were small and that the rates of ozone formation in these air masses were low.

\subsubsection*{Mean AN concentrations}
\begin{table}
\begin{tabular}{cccc}
\hline
Environment   & Mean $RONO_2$ mixing ratio                & $RONO_2$ breakdown & Reference \\
\hline
ITOP?         & mean $\Sigma RONO_2$ 10 ppt               &          & \citep{Reeves2007} \\
???           & mean $\Sigma RONO_2$ 14 ppt               &          & \citep{Roberts1998} \\
remote marine & 2-butyl and 2- and 3-pentyl: 0.8-37 pptv & &\citep{Atherton1989} [Atlas, 1988]\\
urban         & 2-butyl and 2- and 3-pentyl?: approx. 2 ppb & & \citep{Atherton1989} [Hanst et al., 1982] \\
??                         & MeONO2: 1.2 to 43 pptv     &          & \citep{Chuck2002} \\
??                         & EtONO2: 0.8 to 10 pptv     &          & \citep{Chuck2002} \\
remote marine & MeONO2: 0.2 to 50 pptv     & in lit. (2, 8, 9) & \citep{Chuck2002} \\
remote marine & EtONO2: 0.3 to 12.2 pptv   & in lit. (2, 8, 9) & \citep{Chuck2002} \\
Pacific Ocean & MeONO2: 50 pptv            &                   & \citep{Blake2003} \\
airborne      & $\Sigma RONO_2$: approx 1 ppb         & difference between NOy and (NOx+PAN+HNO3) 
                                                        is sample closest to the plume center & \citep{Neuman2012} \\
rural         & mean $\Sigma RONO_2$ (C1-C8): range: 30-630 ppt, mean over the whole period 120 ppt & & \citep{Flocke1998} \\
\hline
\end{tabular}
\end{table}

\subsubsection*{Comparison methodology}
\citep{Sommariva2008}
[The ratios between the alkyl nitrates and their
main alkane precursors can also be plotted against one another, for example against 2-butylnitrate/ n-butane as in Fig. 4. In this way the modelled and the measured concentrations can be compared without taking into account their variation with time. A similar approach was used in previous studies to investigate the sources of alkyl nitrates and the photochemical processing of polluted air masses (e.g., Bertman et al., 1995; Flocke et al., 1998; Roberts et al., 1998; Stroud et al., 2001; Simpson et al., 2003; Reeves et al., 2007). These studies compared plots of the measured ratios against 2-butylnitrate/n-butane to the ratios calculated with a simple sequential reaction model which assumed a single precursors for each alkyl nitrate. Usually, the underestimation of the ratios for short- chain compounds (propylnitrates/propane, ethylni- trate/ethane) was interpreted as an indication of photochemical sources other than the main alkane precursor.]
